---
title: "Almond_ClusterAssign"
author: "Ryan Almond"
date: "2/28/2022"
output: word_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidymodels)
library(cluster) #algorithms for clustering
```

```{r}
trucks <- read_csv("trucks.csv")
```

```{r}
ggplot(trucks, aes(Distance, Speeding)) +
  geom_point(alpha = 0.4) + theme_bw()
```

There appears to be some natural clustering of drivers. Big cluster around the 50 milage distance for drivers not speeding very often and another around the 150-200 mileage range and under 25% speeding. There is another potential cluster for that same distance but a fair amount of drivers speeding more then 50%. A lot more people drive a much shorter distance and typically speed less often.  

```{r}
kmeans_recipe = recipe(~ Speeding + Distance, trucks) 

trucks_dummy = kmeans_recipe %>% 
  step_scale(all_numeric()) %>%
  step_center(all_numeric()) 

trucks_dummy = prep(trucks_dummy, trucks) #prepares the recipe

trucks_cleaned = bake(trucks_dummy, trucks) #applies the recipe and yields a data frame
```

```{r}
set.seed(64)
clusts = 
  tibble(k = 2) %>%
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

clusts
```

```{r}
clusters = 
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>% 
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))
```


```{r}
p1 = 
  ggplot(assignments, aes(x = Distance, y = Speeding)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p1
```

The clusters line up to people who either do not drive very far and people who drive much further. This lines up to what I would image two clusters would be.

```{r}
set.seed(412)
clusts = 
  tibble(k = 1:8) %>%
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

clusts
```

```{r}
clusters = 
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>% 
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))
```

```{r}
p1 = 
  ggplot(assignments, aes(x = Distance, y = Speeding)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p1
```

I think 4 clusters is the right number for what we want to see.  

```{r}
ggplot(clusterings, aes(k, tot.withinss)) +
  geom_line() +
  geom_point()
```

A k value of 3 or 4 seems to be the best for this data. I am going to continue with 4 though based on visualization earlier. 

```{r}
set.seed(64)
clusts = 
  tibble(k = 4) %>%
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

clusts
```

```{r}
clusters = 
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>% 
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))
```

```{r}
p2 = 
  ggplot(assignments, aes(x = Distance, y = Speeding)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p2
```

The clusters that result here are about what I thought when I visualized the relationship early on without clusters present.  